---
title: "vbj-data-pseudobulk"
author: "Genevieve Housman"
date: "April 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pseudobulk data for van Bakel and Jabs - Calculating Pseudobulk

Calculating pseudobulk data for human samples for each cell assignment within the following cell classification schemes:

* stages of differentiation (Stage)
* general unsupervised clusters (Cluster0.05)
* general ad hoc assignments (AdHoc.Assign.thresh0)
* detailed unsupervised clusters (Cluster0.50)
 *osteogenic ad hoc assignments (OstAdHoc.Assign.4gene.thresh0)

```{r, message=FALSE}

#Load libraries
library(Seurat)

##detectCores()
#register(SnowParam(4, "SOCK", progressbar=TRUE))

```

## Prepare data

```{r, eval=FALSE}

#Load data
h.data <- readRDS("./data/h.data.anno.rds")

```

## Function for pseudobulk calculations

Pseudocounts are calculated as sum of raw single-cell UMI counts within each individual-replicate. Using raw counts for pseudobulk data works best for downstream DE analyses.

```{r, eval=FALSE}

runPSEUDO <- function(dataSub, cell.assign, cell.subset) {

  #make pseudocount data using individual+replicate information
  counts <- c()
  metadata <- c()
  labels <- c()
  for (i in unique(dataSub@meta.data$Individual.Replicate)) {
    x.lab <- i
    w <- which(dataSub@meta.data$Individual.Replicate==i)
    if (length(w)>0) {
      x.spp <- dataSub@meta.data$Species[w][1]
      x.col <- dataSub@meta.data$Collection[w][1]
      x.ind <- dataSub@meta.data$Individual[w][1]
      x.rep <- dataSub@meta.data$Replicate[w][1]
      if (length(w)==1) {
        x.cnt <- dataSub@assays$RNA@counts[,w]
      } else {
        x.cnt <- Matrix::rowSums(dataSub@assays$RNA@counts[,w])
      }
      counts <- cbind(counts, x.cnt)
      metadata <- rbind(metadata, c(x.lab, x.spp, x.col, x.ind, x.rep))
      labels <- c(labels, x.lab)
    }
  }
  colnames(counts) <- labels
  rownames(metadata) <- labels
  colnames(metadata) <- c("Individual.Replicate","Species","Collection","Individual","Replicate")
  metadata <- as.data.frame(metadata)

  return(counts)
  
  ##make edgeR object
  #dge <- DGEList(counts)
  #meta_dge <- dge$samples[,c("lib.size","norm.factors")]
  #meta_dge <- cbind(meta_dge, metadata)
  #dge$samples <- meta_dge

}

```

## Scripts for pseudobulk calculations

Run pseudobulk calculations on cluster using one of the following script: pseudobulk.R

* This script performs pseudobulk calculations using the function above.

* The most important argument that needs to be supplemented to the script is as follows:

    + assign = which cell assignments should pairwise tests be performed in (all, stage, cluster0.05, adhoc, cluster0.50, ostadhoc)examines pairwise DE in defined subsets of the data:

* This script was run for the following cell classification systems:

    + Stages of differentiation: Time 0, Time 1, Time 2
    + Unsupervised clustering (res=0.05): iPSC.c1, iPSC.c2, iPSC.c3, MSC.c1, Osteogenic.c1, Osteogenic.c2
    + General ad hoc assignments: iPSCs, MSCs, Osteogenic cells
    + Unsupervised clustering (res=0.50): iPSC.c1, iPSC.c2, iPSC.c3, iPSC.c4, iPSC.c5, iPSC.c6, iPSC.c7, iPSC.c8, MSC.c1, MSC.c2, MSC.c3, MSC.c4, Osteogenic.c1, Osteogenic.c2, Osteogenic.c3, Osteogenic.c4
    + Osteogenic ad hoc assignments: preosteoblasts, osteoblasts, embedding osteoblasts, mineralizing osteoblasts, maturing osteocytes

* This script does not do any of the following steps:

    + gene filtering: kept mitochondrial genes, ribosomal genes, and all other genes filtered
    + TMM normalization of pseudocounts
    + calculationg of cpm (prior.count=0.25)

```{bash, eval=FALSE}

#run pseudobulk.sh/pseudobulk.R on cluster
cd /project2/gilad/ghousman/skeletal-human-chimp/human-skeletal-scRNA/code/

#pseudobulk calculations
sbatch ./pseudobulk.sh stage
sbatch ./pseudobulk.sh cluster0.05
sbatch ./pseudobulk.sh adhoc
sbatch ./pseudobulk.sh cluster0.50
sbatch ./pseudobulk.sh ostadhoc

```

## Do a quick check of DE gene data output.

Similar trends in DE gene numbers seen when replicate 1 samples removed, replicate 2 samples removed, or cells are subsetted to smallest number across species.

```{r, eval=FALSE}

#pseudo <- readRDS("./data/pseudobulk/pseudobulk.stage.Time 0.rds")

pseudo <- read.csv("./data/pseudobulk/pseudobulk.stage.Time 0.csv")
pseudo <- read.csv("./data/pseudobulk/pseudobulk.cluster0.05.iPSC.c1.csv")
pseudo <- read.csv("./data/pseudobulk/pseudobulk.adhoc.iPSC.csv")
pseudo <- read.csv("./data/pseudobulk/pseudobulk.cluster0.50.iPSC.c1.csv")
pseudo <- read.csv("./data/pseudobulk/pseudobulk.ostadhoc.preosteoblast.csv")

dim(pseudo)
head(pseudo)

```
