---
title: "vbj-data-classification"
author: "Genevieve Housman"
date: "April 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pseudobulk data for van Bakel and Jabs - Cell Classification Transfer

Transfer cell classification assignments from Housman et al (2022) PLOS Genetics (GSE181744) to new data.

*stages of differentiation (Stage)
*general unsupervised clusters (Cluster0.05)
*general ad hoc assignments (AdHoc.Assign.thresh0)
*detailed unsupervised clusters (Cluster0.50)
*osteogenic ad hoc assignments (OstAdHoc.Assign.4gene.thresh0)

```{r, message=FALSE}

#Load libraries
library(Seurat)

```

Load new data and data from Housman et al (2022) PLOS Genetics (GSE181744) to new data.

```{r}

#Load new data
h.data <- readRDS("./data/h.data.rds")

#Load old data and isolate just human cells
old_data <- readRDS("./../human-chimp-skeletal-scRNA/data/cellranger-data-full/data.filterC.log.indv-cell.intNo0.reg-tot.assign.rds")
old_data@meta.data$Individual.Replicate <- paste0(old_data@meta.data$Individual,".",old_data@meta.data$Replicate)
old_data.sub <- subset(old_data, subset=Species=="Human")

```

Filter cells to only those examined in Housman et al (2022) PLOS Genetics (GSE181744).

```{r}

#Filter cells
h.data.sub <- subset(h.data, cells=rownames(old_data.sub@meta.data))

#Check filtering process
table(h.data@meta.data$Individual.Replicate)
table(h.data.sub@meta.data$Individual.Replicate)
table(old_data@meta.data$Individual.Replicate[which(old_data@meta.data$Species=="Human")])
table(old_data.sub@meta.data$Individual.Replicate)

```

Transfer cell classification assignments.

```{r}

#Stage
#(already in metadata)

#General unsupervised clusters (Cluster0.05)
h.data.sub@meta.data$Cluster0.05 <- old_data.sub@meta.data$Cluster0.05[which(rownames(old_data.sub@meta.data)==rownames(h.data.sub@meta.data))]

#General ad hoc assignments (AdHoc.Assign.thresh0)
h.data.sub@meta.data$AdHoc.Assign.thresh0 <- old_data.sub@meta.data$AdHoc.Assign.thresh0[which(rownames(old_data.sub@meta.data)==rownames(h.data.sub@meta.data))]

#Detailed unsupervised clusters (Cluster0.50)
h.data.sub@meta.data$Cluster0.50 <- old_data.sub@meta.data$Cluster0.50[which(rownames(old_data.sub@meta.data)==rownames(h.data.sub@meta.data))]

#Osteogenic ad hoc assignments (OstAdHoc.Assign.4gene.thresh0)
h.data.sub@meta.data$OstAdHoc.Assign.4gene.thresh0 <- old_data.sub@meta.data$OstAdHoc.Assign.4gene.thresh0[which(rownames(old_data.sub@meta.data)==rownames(h.data.sub@meta.data))]

#Check transfer process
table(h.data.sub@meta.data$Stage)
table(old_data.sub@meta.data$Stage)

table(h.data.sub@meta.data$Cluster0.05)
table(old_data.sub@meta.data$Cluster0.05)

table(h.data.sub@meta.data$AdHoc.Assign.thresh0)
table(old_data.sub@meta.data$AdHoc.Assign.thresh0)

table(h.data.sub@meta.data$Cluster0.50)
table(old_data.sub@meta.data$Cluster0.50)

table(h.data.sub@meta.data$OstAdHoc.Assign.4gene.thresh0)
table(old_data.sub@meta.data$OstAdHoc.Assign.4gene.thresh0)

```

Save data.

```{r}

saveRDS(h.data.sub, file="./data/h.data.anno.rds")

```
